Simulating High-Dimensional Multivariate Data using the \pkg{Bigsimr} Package

### 2021-05-06 010-background.Rmd
working template
Drafting comparison table in background

\begin{equation}
\rho_P(X,Y) = \frac{E(XY) - E(X)E(Y)}{\left[ \mathrm{Var}(X)\mathrm{Var}(Y)\right]^{1/2}}.
(\#eq:pearson)
\end{equation}

@@

\begin{equation}
(\#eq:convertKendall)
\rho_{P} = \sin \left( \tau \times \frac{\pi}{2} \right), 
\end{equation}

@@

\begin{equation}
(\#eq:convertSpearman)
\rho_P = 2 \times \sin \left( \rho_S \times \frac{\pi}{6} \right).
\end{equation}

@@

\begin{equation}
(\#eq:frechet)
\rho_P^{max} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (U) \right), \quad \rho_P^{min} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (1 - U) \right)
\end{equation}

@@

\begin{equation}
F_{i}^{-1} = \inf\{y:F_{i}(y) \geq u \}.
(\#eq:inverseCDF)
\end{equation}

@@

\begin{equation}
F(x_1, \ldots, x_d) = {\mathbb P}(X_1\leq x_1, \ldots,X_d\leq x_d) = C(F_1(x_1), \ldots, F_d(x_d)), 
(\#eq:copula)
\end{equation}

@@

\begin{equation}
(\#eq:gauss)
F_{{\bf R}}(x_1, \ldots, x_d) = C_{{\bf R}}(\Phi(x_1), \ldots, \Phi(x_d)),
\end{equation}

@@

>>>>

\begin{equation}
\rho_P(X,Y) = \frac{E(XY) - E(X)E(Y)}{\left[ \mathrm{Var}(X)\mathrm{Var}(Y)\right]^{1/2}}.
\label{eq:pearson}
\end{equation}

@@

\begin{equation}
\label{eq:convertKendall}
\rho_{P} = \sin \left( \tau \times \frac{\pi}{2} \right), 
\end{equation}

@@

\begin{equation}
\label{eq:convertSpearman}
\rho_P = 2 \times \sin \left( \rho_S \times \frac{\pi}{6} \right).
\end{equation}

@@

\begin{equation}
\label{eq:frechet}
\rho_P^{max} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (U) \right), \quad \rho_P^{min} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (1 - U) \right)
\end{equation}

@@

\begin{equation}
F_{i}^{-1} = \inf\{y:F_{i}(y) \geq u \}.
\label{eq:inverseCDF}
\end{equation}

@@

\begin{equation}
F(x_1, \ldots, x_d) = {\mathbb P}(X_1\leq x_1, \ldots,X_d\leq x_d) = C(F_1(x_1), \ldots, F_d(x_d)), 
\label{eq:copula}
\end{equation}

@@

\begin{equation}
\label{eq:gauss}
F_{{\bf R}}(x_1, \ldots, x_d) = C_{{\bf R}}(\Phi(x_1), \ldots, \Phi(x_d)),
\end{equation}

@@

## Relevant Multivariate Simulation software Comparison

```{r tab:compare-table}
knitr::kable(
           head(mtcars[, 1:8], 10),  booktabs = TRUE, format = "latex",
                 linesep = c("", "", "\\addlinespace"),
           caption = 'A table of the first 10 rows of the mtcars data.'
  )
```

<!-- 
| Software | Designed for | Scales to Large P | Hetereogenous Margins | Flexible Dependence |
|----------|--------------|-------------------|-----------------------|---------------------|
|`copula`|              |                   |                       |                     |
|`genord`|              |                   |                       |                     |
|`mvnfast`|              |                   |                       |                     |

your comment -->

### 2021-05-06 article_bigsimr.tex
Drafting comparison table in background

@@

\begin{equation}
    \mathrm{E}\left[Y_i Y_j\right] = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} F_i^{-1}\left[\Phi(z_i)\right] F_j^{-1}\left[\Phi(z_j)\right] \phi(z_i, z_j, \rho_z) dz_i dz_j,
    (\#eq:pearsonIntegralRelation)
\end{equation}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> bounds <- bs$cor_bounds(margins)
R> bounds$lower
\end{CodeInput}
\begin{CodeOutput}
        [,1]    [,2]
[1,]  1.0000 -0.8745
[2,] -0.8745  1.0000
\end{CodeOutput}
\begin{CodeInput}
R> bounds$upper
\end{CodeInput}
\begin{CodeOutput}
       [,1]   [,2]
[1,] 1.0000 0.8748
[2,] 0.8748 1.0000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> bs$cor_bounds(margins[1], margins[2], bs$Pearson, n_samples = 1e6)
\end{CodeInput}
\begin{CodeOutput}
Julia Object of type NamedTuple{(:lower, :upper),Tuple{Float64,Float64}}.
(lower = -0.8799819883897544, upper = 0.8798667127065275)
\end{CodeOutput}
\end{CodeChunk}

@@

\caption{\label{tab:ch040-BiError}Average abolute error in matching the target dependency across the entire range of possible correlations for each bivariate marginal.}
\centering
\begin{tabular}[t]{lllr}
\toprule
No. of random vectors & Correlation type & Distribution & Mean abs. error\\
\midrule
1000 & Pearson & norm & 0.0168\\
1000 & Pearson & gamma & 0.0159\\
1000 & Pearson & nbinom & 0.0188\\
\addlinespace
1000 & Spearman & norm & 0.0203\\
1000 & Spearman & gamma & 0.0180\\
1000 & Spearman & nbinom & 0.0165\\
\addlinespace
1000 & Kendall & norm & 0.0118\\
1000 & Kendall & gamma & 0.0099\\
1000 & Kendall & nbinom & 0.0089\\
\addlinespace
10000 & Pearson & norm & 0.0048\\
10000 & Pearson & gamma & 0.0055\\
10000 & Pearson & nbinom & 0.0066\\
\addlinespace
10000 & Spearman & norm & 0.0058\\
10000 & Spearman & gamma & 0.0054\\
10000 & Spearman & nbinom & 0.0060\\
\addlinespace
10000 & Kendall & norm & 0.0034\\
10000 & Kendall & gamma & 0.0031\\
10000 & Kendall & nbinom & 0.0037\\
\addlinespace
1e+05 & Pearson & norm & 0.0016\\
1e+05 & Pearson & gamma & 0.0016\\
1e+05 & Pearson & nbinom & 0.0032\\
\addlinespace
1e+05 & Spearman & norm & 0.0018\\
1e+05 & Spearman & gamma & 0.0019\\
1e+05 & Spearman & nbinom & 0.0013\\
\addlinespace
1e+05 & Kendall & norm & 0.0011\\
1e+05 & Kendall & gamma & 0.0010\\
1e+05 & Kendall & nbinom & 0.0010\\
\bottomrule
\end{tabular}
\end{table}

@@

|Dimension |Correlation type | Total Time (Seconds)|
|:---------|:----------------|--------------------:|
|100       |Pearson          |                0.080|
|100       |Spearman         |                0.036|
|100       |Kendall          |                0.116|
|250       |Pearson          |                0.286|
|250       |Spearman         |                0.371|
|250       |Kendall          |                0.366|
|500       |Pearson          |                2.932|
|500       |Spearman         |                0.176|
|500       |Kendall          |                2.183|

@@

where \({\bf Y}^{(b)}\) is the \(b^{th}\) simulated vector from a total
of \(B\) simulation replicates and \(I(\cdot)\) is the indicator
function. For example, we can estimate from our \(B=10,000\) simulated
vectors that the probability of all genes expressing (i.e.,
\({\bf y}_i \geq 1, \forall \; i\)) is \(0.5979\).

\begin{CodeChunk}
\begin{CodeInput}
R> d <- ncol(sim_nbinom)
R> B <- nrow(sim_nbinom)
R> y0 <- 1
R> pHat <- mean(apply(sim_nbinom, 1, function(Y) {
+   all(Y >= y0)
+ }))
R> pHat
\end{CodeInput}
\begin{CodeOutput}
[1] 0.5979
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summary(frobenius_loss)
\end{CodeInput}
\begin{CodeOutput}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    691     829     890     925     986    1224 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{equation}
(Y_1, Y_2)^\top \stackrel{d}{=} \left(X_1, e^{X_2}\right)^\top,
(\#eq:kram1)
\end{equation}

@@

\begin{equation}
\boldsymbol \Sigma = 
\begin{bmatrix}
\sigma_1^2 & \rho \sigma_1\sigma_2\\
\rho \sigma_1\sigma_2 & \sigma_2^2
\end{bmatrix}
(\#eq:kram2)
\end{equation}

@@

\begin{equation}
\boldsymbol R = 
\left[
\begin{array}{cc}
1 & \rho \\
\rho & 1
\end{array}
\right].
(\#eq:kram3)
\end{equation}

@@

\begin{equation}
F_1^{-1}(u) = \mu_1+\sigma_1 \Phi^{-1}(u)\,\,\, \mbox{and} \,\,\, F_2^{-1}(u) = e^{\mu_2+\sigma_2 \Phi^{-1}(u)}, 
(\#eq:kram4)
\end{equation}

@@

\begin{equation}
F_1^{-1}(U_1) = \mu_1+\sigma_1 \Phi^{-1}(\Phi(Z_1)) =  \mu_1+\sigma_1 Z_1 \,\,\, \mbox{and} \,\,\, F_2^{-1}(U_2) = e^{\mu_2+\sigma_2 \Phi^{-1}(\Phi(Z_2))} = e^{\mu_2+\sigma_2 Z_2}, 
(\#eq:kram5)
\end{equation}

where
\((X_1, X_2)^\top = (\mu_1+\sigma_1 Z_1, \mu_2+\sigma_2 Z_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)\).
Consequently, the vector \((Y_1, Y_2)^\top\) obtained in Step 3 of the
algorithm has the structure provided by (\ref{kram1}).

@@

\noindent Next, we provide the exact covariance structure of the random
vector \((Y_1, Y_2)^\top\) given by (\ref{kram1}), and relate the
correlation of \(Y_1\) and \(Y_2\) to \(\rho\), which is the correlation
of the normal variables \(Z_1\), \(Z_2\) (and also \(X_1\) and \(X_2\)).
A straightforward albeit somewhat tedious algebra produces the following
result.

\begin{lemma}
Let ${\bf Y} = (Y_1, Y_2)^\top$ admit the stochastic representation (\ref{kram1}), where ${\bf X} = (X_1, X_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)$ with $\boldsymbol \mu$ and $\boldsymbol \Sigma$ as above. Then the mean vector and the variance-covariance matrix of ${\bf Y}$ are given by

\begin{equation}
\boldsymbol \mu_{{\bf Y}} = \left(\mu_1, e^{\mu_2+\sigma_2^2/2}\right)^\top
(\#eq:kram6)
\end{equation}

@@

\begin{equation}
\boldsymbol \Sigma = 
\begin{bmatrix}
\sigma_1^2 & \rho \sigma_1\sigma_2  e^{\mu_2+\sigma_2^2/2}  \\
\rho \sigma_1\sigma_2 e^{\mu_2+\sigma_2^2/2} & \left[ e^{\mu_2+\sigma_2^2/2}\right]^2 \left[e^{\sigma_2^2} -1 \right],
\end{bmatrix}
(\#eq:kram7)
\end{equation}

@@

\begin{equation}
\mbox{Cov}(Y_1, Y_2) = \mathbb E(Y_1 Y_2) - \mathbb E(Y_1)\mathbb E(Y_2), 
(\#eq:covy1y2)
\end{equation}

@@

\begin{equation}
\mathbb E(Y_1 Y_2)  =  \mathbb E\left(X_1 e^{X_2}\right)  =  \mathbb E \left\{ \mathbb E \left(X_1 e^{X_2} | X_1 \right) \right\} = \mathbb E \left\{ X_1 \mathbb E \left(e^{X_2} | X_1 \right) \right\}.
(\#eq:kre1)
\end{equation}

Further, by using the fact that the conditional distribution of $X_2$ given $X_1=x_1$ is normal with mean $\mathbb E(X_2|X_1=x_1) = \mu_2+\rho\sigma_2(x_1-\mu_1)/\sigma_1$ and variance $\sigma_2^2(1-\rho^2)$, one can relate the inner expectation on the far right in (\ref{kre1}) to the moment generating function $M(t)$ of this conditional distribution evaluated at $t=1$, leading to

\begin{equation}
E \left(e^{X_2} | X_1 \right) = e^{ \mu_2+\rho\sigma_2(x_1-\mu_1)/\sigma_1 + \sigma_2^2(1-\rho^2)/2},
(\#eq:kre2)
\end{equation}

@@

\begin{equation}
\mathbb E(Y_1 Y_2)  =  e^{ \mu_2+ \frac{1}{2} \sigma_2^2(1-\rho^2)} \mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\}.
(\#eq:kre3)
\end{equation}

Since $X_1$ is normal with mean $\mu_1$ and variance $\sigma_1^2$, the expectation in (\ref{kre3}) becomes

\begin{equation}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = \int_{-\infty}^\infty x_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (x_1-\mu_1)} \frac{1}{\sqrt{2\pi}\sigma_1} e^{-\frac{1}{2\sigma_1^2}(x_1-\mu_1)^2}.
(\#eq:kre4)
\end{equation}

@@

\begin{equation}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = e^{\frac{1}{2}\sigma_2^2\rho^2} \int_{-\infty}^\infty (u+\mu_1)  \frac{1}{\sqrt{2\pi}\sigma_1} e^{-\frac{1}{2\sigma_1^2}(u-\rho\sigma_1\sigma_2)^2}.
(\#eq:kre5)
\end{equation}

Since the integral in (\ref{kre5}) is the expectation $\mathbb E(X+\mu_1)$, where $X$ is normal with mean $\rho\sigma_1\sigma_2$ and variance $\sigma_1^2$, we conclude that

\begin{equation}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = e^{\frac{1}{2}\sigma_2^2\rho^2} (\rho\sigma_1\sigma_2 +\mu_1),
(\#eq:kre6)
\end{equation}

which, in view of (\ref{kre3}), leads to

\begin{equation}
\mathbb E (Y_1Y_2) = e^{\mu_2 + \frac{1}{2}\sigma_2^2} (\rho\sigma_1\sigma_2 +\mu_1).
(\#eq:kre7)
\end{equation}

Finally, (\ref{covy1y2}), along with the expressions for the means of $Y_1$ and $Y_2$, produce the covariance of $Y_1$ and $Y_2$:

\begin{equation}
\mbox{Cov}(Y_1, Y_2) = e^{\mu_2 + \frac{1}{2}\sigma_2^2} (\rho\sigma_1\sigma_2 +\mu_1) - \mu_1 e^{\mu_2 + \frac{1}{2}\sigma_2^2} = \rho\sigma_1\sigma_2 e^{\mu_2 + \frac{1}{2}\sigma_2^2}.
(\#eq:kre8)
\end{equation}

@@

\begin{equation}
\rho_{Y_1, Y_2} = \rho \frac{\sigma_2}{\sqrt{e^{\sigma_2^2} -1}},
(\#eq:kram8)
\end{equation}

where \(\rho\) is the correlation of \(X_1\) and \(X_2\). The above
result is useful when studying the possible range of correlation of
\(Y_1\) and \(Y_2\) in the above example. It can be shown that the
factor on the far right in (\ref{kram8}) is a monotonically decreasing
function of \(\sigma_2\) on \((0,\infty)\), with the limits of 1 and 0
at zero and infinity, respectively. Thus, in principle, the range of
correlation of \(Y_1\) and \(Y_2\) is the same as that of \(X_1\) and
\(X_2\), as the factor on the far right in (\ref{kram8}) can be made
arbitrarily close to 1. However, by changing this factor we may affect
the marginal distributions of \(Y_1\) and \(Y_2\). It can be shown that
if the marginal distributions of \(Y_1\) and \(Y_2\) are fixed, then the
relation (\ref{kram8}) becomes

\begin{equation}
\rho_{Y_1, Y_2} = \rho \sqrt{\frac{\log(1+c_2^2)}{c_2^2}},
(\#eq:kram9)
\end{equation}

@@

\begin{equation}
- \sqrt{\frac{\log(1+c_2^2)}{c_2^2}} \leq \rho_{Y_1, Y_2} \leq \sqrt{\frac{\log(1+c_2^2)}{c_2^2}}. 
(\#eq:kram10)
\end{equation}

@@

\begin{equation}
\frac{\hat{\sigma}_2}{\sqrt{e^{\hat{\sigma}_2^2} -1}} = 0.881.
(\#eq:kram11)
\end{equation}

>>>>

\hypertarget{relevant-multivariate-simulation-software-comparison}{%
\subsection{Relevant Multivariate Simulation software
Comparison}\label{relevant-multivariate-simulation-software-comparison}}

\caption{\label{tab:tab:compare-table}A table of the first 10 rows of the mtcars data.}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
  & mpg & cyl & disp & hp & drat & wt & qsec & vs\\
\midrule
Mazda RX4 & 21.0 & 6 & 160.0 & 110 & 3.90 & 2.620 & 16.46 & 0\\
Mazda RX4 Wag & 21.0 & 6 & 160.0 & 110 & 3.90 & 2.875 & 17.02 & 0\\
Datsun 710 & 22.8 & 4 & 108.0 & 93 & 3.85 & 2.320 & 18.61 & 1\\
\addlinespace
Hornet 4 Drive & 21.4 & 6 & 258.0 & 110 & 3.08 & 3.215 & 19.44 & 1\\
Hornet Sportabout & 18.7 & 8 & 360.0 & 175 & 3.15 & 3.440 & 17.02 & 0\\
Valiant & 18.1 & 6 & 225.0 & 105 & 2.76 & 3.460 & 20.22 & 1\\
\addlinespace
Duster 360 & 14.3 & 8 & 360.0 & 245 & 3.21 & 3.570 & 15.84 & 0\\
Merc 240D & 24.4 & 4 & 146.7 & 62 & 3.69 & 3.190 & 20.00 & 1\\
Merc 230 & 22.8 & 4 & 140.8 & 95 & 3.92 & 3.150 & 22.90 & 1\\
\addlinespace
Merc 280 & 19.2 & 6 & 167.6 & 123 & 3.92 & 3.440 & 18.30 & 1\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{equation}
    \mathrm{E}\left[Y_i Y_j\right] = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} F_i^{-1}\left[\Phi(z_i)\right] F_j^{-1}\left[\Phi(z_j)\right] \phi(z_i, z_j, \rho_z) dz_i dz_j,
    \label{eq:pearsonIntegralRelation}
\end{equation}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> bounds <- bs$cor_bounds(margins)
R> bounds$lower
\end{CodeInput}
\begin{CodeOutput}
       [,1]   [,2]
[1,]  1.000 -0.878
[2,] -0.878  1.000
\end{CodeOutput}
\begin{CodeInput}
R> bounds$upper
\end{CodeInput}
\begin{CodeOutput}
      [,1]  [,2]
[1,] 1.000 0.879
[2,] 0.879 1.000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> bs$cor_bounds(margins[1], margins[2], bs$Pearson, n_samples = 1e6)
\end{CodeInput}
\begin{CodeOutput}
Julia Object of type NamedTuple{(:lower, :upper),Tuple{Float64,Float64}}.
(lower = -0.8801373451421908, upper = 0.8801357032571077)
\end{CodeOutput}
\end{CodeChunk}

@@

\caption{\label{tab:ch040-BiError}Average abolute error in matching the target dependency across the entire range of possible correlations for each bivariate marginal.}
\centering
\begin{tabular}[t]{lllr}
\toprule
No. of random vectors & Correlation type & Distribution & Mean abs. error\\
\midrule
1000 & Pearson & norm & 0.0156\\
1000 & Pearson & gamma & 0.0159\\
1000 & Pearson & nbinom & 0.0162\\
\addlinespace
1000 & Spearman & norm & 0.0175\\
1000 & Spearman & gamma & 0.0182\\
1000 & Spearman & nbinom & 0.0151\\
\addlinespace
1000 & Kendall & norm & 0.0130\\
1000 & Kendall & gamma & 0.0115\\
1000 & Kendall & nbinom & 0.0126\\
\addlinespace
10000 & Pearson & norm & 0.0060\\
10000 & Pearson & gamma & 0.0058\\
10000 & Pearson & nbinom & 0.0058\\
\addlinespace
10000 & Spearman & norm & 0.0062\\
10000 & Spearman & gamma & 0.0056\\
10000 & Spearman & nbinom & 0.0049\\
\addlinespace
10000 & Kendall & norm & 0.0033\\
10000 & Kendall & gamma & 0.0038\\
10000 & Kendall & nbinom & 0.0033\\
\addlinespace
1e+05 & Pearson & norm & 0.0018\\
1e+05 & Pearson & gamma & 0.0017\\
1e+05 & Pearson & nbinom & 0.0030\\
\addlinespace
1e+05 & Spearman & norm & 0.0017\\
1e+05 & Spearman & gamma & 0.0016\\
1e+05 & Spearman & nbinom & 0.0015\\
\addlinespace
1e+05 & Kendall & norm & 0.0010\\
1e+05 & Kendall & gamma & 0.0011\\
1e+05 & Kendall & nbinom & 0.0012\\
\bottomrule
\end{tabular}
\end{table}

@@

|Dimension |Correlation type | Total Time (Seconds)|
|:---------|:----------------|--------------------:|
|100       |Pearson          |                0.081|
|100       |Spearman         |                0.027|
|100       |Kendall          |                0.075|
|250       |Pearson          |                0.378|
|250       |Spearman         |                0.062|
|250       |Kendall          |                0.375|
|500       |Pearson          |                1.425|
|500       |Spearman         |                0.135|
|500       |Kendall          |                1.495|

@@

where \({\bf Y}^{(b)}\) is the \(b^{th}\) simulated vector from a total
of \(B\) simulation replicates and \(I(\cdot)\) is the indicator
function. For example, we can estimate from our \(B=10,000\) simulated
vectors that the probability of all genes expressing (i.e.,
\({\bf y}_i \geq 1, \forall \; i\)) is \(0.6015\).

\begin{CodeChunk}
\begin{CodeInput}
R> d <- ncol(sim_nbinom)
R> B <- nrow(sim_nbinom)
R> y0 <- 1
R> pHat <- mean(apply(sim_nbinom, 1, function(Y) {
+   all(Y >= y0)
+ }))
R> pHat
\end{CodeInput}
\begin{CodeOutput}
[1] 0.6015
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summary(frobenius_loss)
\end{CodeInput}
\begin{CodeOutput}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    675     813     974     998    1080    1851 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{equation}
\label{eq:kram1}
(Y_1, Y_2)^\top \stackrel{d}{=} \left(X_1, e^{X_2}\right)^\top,
\end{equation}

@@

\begin{equation}
\label{eq:kram2}
\boldsymbol \Sigma = 
\begin{bmatrix}
\sigma_1^2 & \rho \sigma_1\sigma_2\\
\rho \sigma_1\sigma_2 & \sigma_2^2
\end{bmatrix}
\end{equation}

@@

\begin{equation}
\label{eq:kram3}
\boldsymbol R = 
\left[
\begin{array}{cc}
1 & \rho \\
\rho & 1
\end{array}
\right].
\end{equation}

@@

\begin{equation}
\label{eq:kram4}
F_1^{-1}(u) = \mu_1+\sigma_1 \Phi^{-1}(u)\,\,\, \mbox{and} \,\,\, F_2^{-1}(u) = e^{\mu_2+\sigma_2 \Phi^{-1}(u)}, 
\end{equation}

@@

\begin{equation}
\label{eq:kram5}
F_1^{-1}(U_1) = \mu_1+\sigma_1 \Phi^{-1}(\Phi(Z_1)) =  \mu_1+\sigma_1 Z_1 \,\,\, \mbox{and} \,\,\, F_2^{-1}(U_2) = e^{\mu_2+\sigma_2 \Phi^{-1}(\Phi(Z_2))} = e^{\mu_2+\sigma_2 Z_2}, 
\end{equation}

where
\((X_1, X_2)^\top = (\mu_1+\sigma_1 Z_1, \mu_2+\sigma_2 Z_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)\).
Consequently, the vector \((Y_1, Y_2)^\top\) obtained in Step 3 of the
algorithm has the structure provided by (\ref{eq:kram1}).

@@

\noindent Next, we provide the exact covariance structure of the random
vector \((Y_1, Y_2)^\top\) given by (\ref{eq:kram1}), and relate the
correlation of \(Y_1\) and \(Y_2\) to \(\rho\), which is the correlation
of the normal variables \(Z_1\), \(Z_2\) (and also \(X_1\) and \(X_2\)).
A straightforward albeit somewhat tedious algebra produces the following
result.

\begin{lemma}
Let ${\bf Y} = (Y_1, Y_2)^\top$ admit the stochastic representation (\ref{eq:kram1}), where ${\bf X} = (X_1, X_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)$ with $\boldsymbol \mu$ and $\boldsymbol \Sigma$ as above. Then the mean vector and the variance-covariance matrix of ${\bf Y}$ are given by

\begin{equation}
\label{eq:kram6}
\boldsymbol \mu_{{\bf Y}} = \left(\mu_1, e^{\mu_2+\sigma_2^2/2}\right)^\top
\end{equation}

@@

\begin{equation}
\label{eq:kram7}
\boldsymbol \Sigma = 
\begin{bmatrix}
\sigma_1^2 & \rho \sigma_1\sigma_2  e^{\mu_2+\sigma_2^2/2}  \\
\rho \sigma_1\sigma_2 e^{\mu_2+\sigma_2^2/2} & \left[ e^{\mu_2+\sigma_2^2/2}\right]^2 \left[e^{\sigma_2^2} -1 \right],
\end{bmatrix}
\end{equation}

@@

\begin{equation}
\label{eq:covy1y2}
\mbox{Cov}(Y_1, Y_2) = \mathbb E(Y_1 Y_2) - \mathbb E(Y_1)\mathbb E(Y_2), 
\end{equation}

@@

\begin{equation}
\label{eq:kre1}
\mathbb E(Y_1 Y_2)  =  \mathbb E\left(X_1 e^{X_2}\right)  =  \mathbb E \left\{ \mathbb E \left(X_1 e^{X_2} | X_1 \right) \right\} = \mathbb E \left\{ X_1 \mathbb E \left(e^{X_2} | X_1 \right) \right\}.
\end{equation}

Further, by using the fact that the conditional distribution of $X_2$ given $X_1=x_1$ is normal with mean $\mathbb E(X_2|X_1=x_1) = \mu_2+\rho\sigma_2(x_1-\mu_1)/\sigma_1$ and variance $\sigma_2^2(1-\rho^2)$, one can relate the inner expectation on the far right in (\ref{eq:kre1}) to the moment generating function $M(t)$ of this conditional distribution evaluated at $t=1$, leading to

\begin{equation}
\label{eq:kre2}
E \left(e^{X_2} | X_1 \right) = e^{ \mu_2+\rho\sigma_2(x_1-\mu_1)/\sigma_1 + \sigma_2^2(1-\rho^2)/2},
\end{equation}

@@

\begin{equation}
\label{eq:kre3}
\mathbb E(Y_1 Y_2)  =  e^{ \mu_2+ \frac{1}{2} \sigma_2^2(1-\rho^2)} \mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\}.
\end{equation}

Since $X_1$ is normal with mean $\mu_1$ and variance $\sigma_1^2$, the expectation in (\ref{eq:kre3}) becomes

\begin{equation}
\label{eq:kre4}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = \int_{-\infty}^\infty x_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (x_1-\mu_1)} \frac{1}{\sqrt{2\pi}\sigma_1} e^{-\frac{1}{2\sigma_1^2}(x_1-\mu_1)^2}.
\end{equation}

@@

\begin{equation}
\label{eq:kre5}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = e^{\frac{1}{2}\sigma_2^2\rho^2} \int_{-\infty}^\infty (u+\mu_1)  \frac{1}{\sqrt{2\pi}\sigma_1} e^{-\frac{1}{2\sigma_1^2}(u-\rho\sigma_1\sigma_2)^2}.
\end{equation}

Since the integral in (\ref{eq:kre5}) is the expectation $\mathbb E(X+\mu_1)$, where $X$ is normal with mean $\rho\sigma_1\sigma_2$ and variance $\sigma_1^2$, we conclude that

\begin{equation}
\label{eq:kre6}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = e^{\frac{1}{2}\sigma_2^2\rho^2} (\rho\sigma_1\sigma_2 +\mu_1),
\end{equation}

which, in view of (\ref{eq:kre3}), leads to

\begin{equation}
\label{eq:kre7}
\mathbb E (Y_1Y_2) = e^{\mu_2 + \frac{1}{2}\sigma_2^2} (\rho\sigma_1\sigma_2 +\mu_1).
\end{equation}

Finally, (\ref{eq:covy1y2}), along with the expressions for the means of $Y_1$ and $Y_2$, produce the covariance of $Y_1$ and $Y_2$:

\begin{equation}
\label{eq:kre8}
\mbox{Cov}(Y_1, Y_2) = e^{\mu_2 + \frac{1}{2}\sigma_2^2} (\rho\sigma_1\sigma_2 +\mu_1) - \mu_1 e^{\mu_2 + \frac{1}{2}\sigma_2^2} = \rho\sigma_1\sigma_2 e^{\mu_2 + \frac{1}{2}\sigma_2^2}.
\end{equation}

@@

\begin{equation}
\label{eq:kram8}
\rho_{Y_1, Y_2} = \rho \frac{\sigma_2}{\sqrt{e^{\sigma_2^2} -1}},
\end{equation}

where \(\rho\) is the correlation of \(X_1\) and \(X_2\). The above
result is useful when studying the possible range of correlation of
\(Y_1\) and \(Y_2\) in the above example. It can be shown that the
factor on the far right in (\ref{eq:kram8}) is a monotonically
decreasing function of \(\sigma_2\) on \((0,\infty)\), with the limits
of 1 and 0 at zero and infinity, respectively. Thus, in principle, the
range of correlation of \(Y_1\) and \(Y_2\) is the same as that of
\(X_1\) and \(X_2\), as the factor on the far right in (\ref{eq:kram8})
can be made arbitrarily close to 1. However, by changing this factor we
may affect the marginal distributions of \(Y_1\) and \(Y_2\). It can be
shown that if the marginal distributions of \(Y_1\) and \(Y_2\) are
fixed, then the relation (\ref{eq:kram8}) becomes

\begin{equation}
\label{eq:kram9}
\rho_{Y_1, Y_2} = \rho \sqrt{\frac{\log(1+c_2^2)}{c_2^2}},
\end{equation}

@@

\begin{equation}
\label{eq:kram10}
- \sqrt{\frac{\log(1+c_2^2)}{c_2^2}} \leq \rho_{Y_1, Y_2} \leq \sqrt{\frac{\log(1+c_2^2)}{c_2^2}}. 
\end{equation}

@@

\begin{equation}
\label{eq:kram11}
\frac{\hat{\sigma}_2}{\sqrt{e^{\hat{\sigma}_2^2} -1}} = 0.881.
\end{equation}
